---
title: "Assignment 2"
output: pdf_document
date: '2022-03-11'
---

---
title: "ST306 assignment 2"
output: pdf
date: '2022-02-28'
format: 'rmarkdown'
---
Q1:
(a) True, the normally distributed residuals reveal that there is no trending or seasonal information in the data
which has not been explained by the forecasting model
(b) False , may be a good forecasting model but not always true , a good forecast needs to have residuals 
of zero mean , constant variance and follow a normal distribution.
(c) False , there is no best measure for forecasting accuracy 
(d) False , more complex model does not ensure a more accurate model, if a model dosent work well we should first identify why it doesnt and go from there.
(e) True , the objective of a model is to make correct forcasting for new data.

Q2:
```{r}
library(fpp3)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(tsibble)
library(tsibbledata)
library(stringr)
library(insight)
library(seasonal)
library(tinytex===)
set.seed(12357695)
myseries <- aus_retail %>%
  filter(`Series ID` == sample(aus_retail$`Series ID`,1))


```
(a)
```{r}
myseries_train <- myseries %>% as_tsibble() %>%
  filter(year(Month) < 2011)

 
```
(b)
```{r}
library(forecast)
library(ggplot2)
autoplot(myseries, Turnover) +
  autolayer(myseries_train, Turnover, colour = "red")

 
```
(c)
```{r}


fit <- myseries_train %>% 
  model(SNAIVE(Turnover))
  
  
```

(d)
```{r}
fit %>% gg_tsresiduals()
```
(e)
```{r}
fc <- fit %>%
  forecast(new_data = anti_join(myseries, myseries_train))
fc %>% autoplot(myseries)
```
(f)
```{r}
fit %>% accuracy()
fc %>% accuracy(myseries)
```
(g)
The training data was split from the tail of the data and thus the accuracy measures are highly sensitive to the training /test split , given that there is significant grows in the trend and variability of the data.
Q3
(a)
```{r}

autoplot(canadian_gas)
gg_subseries(canadian_gas)
gg_season(canadian_gas)
```
(b)
```{r}
fit_stl <- stl(canadian_gas, s.window = 13, robust = TRUE)
autoplot(fit_stl) + xlab("Month") + ggtitle("STL Decomposition of Canadian Gas Production")


```

(c)
The seasonal shape starts off narrow and then as it gets closer to the 1990s the variance increases and takes a  trumpet shape and then as it moves towards the 2000s is slowly narrows and the variance decreases.

(d)
```{r}
autoplot(canadian_gas, series="Data") +
  autolayer(trendcycle(fit_stl), series="Trend") +
  autolayer(seasadj(fit_stl), series="Seasonally Adjusted") +
  xlab("Year") + ylab("New orders index") +
  ggtitle("STL Decomposition of Canadian Gas Production") +
  scale_colour_manual(values=c("gray","blue","red"),
             breaks=c("Data","Seasonally Adjusted","Trend"))
```
(e)
```{r}


fit_seats <- canadian_gas %>% model(X_13ARIMA_SEATS(Volume ~ seats()))
#p1 <- autoplot(fit_seats)
fit_x11 <- canadian_gas %>% model(X_13ARIMA_SEATS(Volume ~ x11()))

```


